---
title: "PA1_template.Rmd"
author: "ME!"
date: "1/3/2022"
output: html_document
---

```{r}
originalData <- read.csv("activity.csv",header=TRUE)
datalessNA <- na.omit(originalData)
```

## 1. What is mean total number of steps taken per day?

## 1A. Calculate the total number of steps taken per day


```{r, echo = TRUE}
steps_each_day <- aggregate(datalessNA$steps, by= list(datalessNA$date), FUN="sum")
steps_each_day
```

## 1B. Make a histogram of the total number of steps taken each day

```{r plot}
hist(steps_each_day$x, col="blue", xlab = "Number of steps per day",
     main = "Total number of steps taken each day", breaks = 30)
```

```{r plot histogram}
plot(hist(steps_each_day$x, col="blue", xlab = "Number of steps per day",
     main = "Total number of steps taken each day", breaks = 30))
```

## 1C. Calculate and report the mean and median of the total number of steps taken per day

```{r mean steps}
mean_steps <- mean(steps_each_day[,2])
mean_steps
```

```{r median steps}
median_steps <- median(steps_each_day[,2])
median_steps
```

## 2. What is the average daily activity pattern?

## 2A. Make a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)

```{r mean_each_day}
mean_each_day <- aggregate(datalessNA$steps, by = list(Interval = datalessNA$interval), FUN = "mean")
```

```{r plot time series}
plot(mean_each_day$Interval, mean_each_day$x, type = "l", xlab = "5-min intervals",
     ylab = "Average number of steps taken", main = " Average daily activity pattern")
```

## 2B. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r interval with max number of steps}
interval_row <- which.max(mean_each_day$x)
max_interval <- mean_each_day[interval_row, 1]
max_interval
```

## 3. Note that there are a number of days/intervals where there are missing values (coded as red NA). The presence of missing days may introduce bias into some calculations or summaries of the data.

## 3A. Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)

```{r number of NAs, echo = TRUE}
number_of_NAs <- length(which(is.na(originalData$steps)))
number_of_NAs
```

## 3B. Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.

## 3C.Create a new dataset that is equal to the original dataset but with the missing data filled in.

````{r fill NAs with data and create new dataset}
library(Hmisc)
originalData_filled <- originalData 
originalData_filled$steps <- impute(originalData$steps, fun = mean)
```

## 3D. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment? What is the impact of imputing missing data on the estimates of the total daily number of steps?

```{r plot histogram with NAs filled}
steps_each_day_noNAs <- aggregate(originalData_filled$steps, by = list(originalData_filled$date), FUN="sum")
hist(steps_each_day_noNAs$x, col="green", xlab = "Number of steps per day",
     main = "Total number of steps taken each day", breaks = 30)
```

```{r mean steps with NAs filled, echo = TRUE}
mean_steps_NAsfilled <- mean(steps_each_day_noNAs[,2])
mean_steps_NAsfilled
```

```{r median_steps, echo = TRUE}
median_steps_NAsfilled <- median(steps_each_day_noNAs[,2])
median_steps_NAsfilled
```

## 5. Are there differences in activity patterns between weekdays and weekends?

## 5A. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r create dataset indicating weekday and weekend}
originalData_filled$date <- as.Date(originalData_filled$date)
originalData_filled$weekday <- weekdays(originalData_filled$date)
originalData_filled$weekend <- ifelse(originalData_filled$weekday == "Saturday" | originalData_filled$weekday == "Sunday", "Weekend", "Weekday")
```

## 5B. Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.

```{r plot by day type dataset}
day_type <- aggregate(steps ~ interval + weekend, data = originalData_filled, mean)
library(ggplot2)
ggplot(day_type, aes(interval, steps)) + geom_line() + facet_grid(weekend ~ .) +
        xlab("5-minute intervals") + 
        ylab("Avarage number of steps taken") + 
        ggtitle ("Weekdays & Weekends activity pattern")
```

